Capstone Movielens Report

### 2025-05-02 reports/0ld/capstone-movielens-report.draft3.tex
Draft version 4 of the RMD report compiled.

\title{Capstone Movielens Report}
\author{Azamat Kurbanaev}
\date{2025-02-21}

@@

@@

\includegraphics{capstone-movielens-report.draft1_files/figure-latex/unnamed-chunk-10-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft1_files/figure-latex/unnamed-chunk-11-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft1_files/figure-latex/unnamed-chunk-14-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft1_files/figure-latex/unnamed-chunk-15-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft1_files/figure-latex/unnamed-chunk-18-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft1_files/figure-latex/unnamed-chunk-19-1.pdf}

@@

\includegraphics[width=0.4\linewidth]{capstone-movielens-report.draft1_files/figure-latex/sparsity-of-movie-recs-1}

@@

\includegraphics{capstone-movielens-report.draft1_files/figure-latex/movie-id-and-user-hists-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft1_files/figure-latex/unnamed-chunk-32-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft1_files/figure-latex/unnamed-chunk-33-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft1_files/figure-latex/unnamed-chunk-39-1.pdf}
Now we can determine the minimal \emph{RMSE}:

@@

\includegraphics{capstone-movielens-report.draft1_files/figure-latex/unnamed-chunk-45-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft1_files/figure-latex/unnamed-chunk-46-1.pdf}
Below are worst and best ratings categories:

@@

\includegraphics{capstone-movielens-report.draft1_files/figure-latex/unnamed-chunk-48-1.pdf}

@@

\[
Y_{i,j} = \mu + \alpha_i + \beta_j + f(d_{i,j}) + \sum_{k=1}^K x_{i,j}^k \gamma_k + \varepsilon_{i,j}
\]

with \(x^k_{i,j} = 1\) if \(g_{i,j}\) is genre \(k\).

>>>>

\title{Capstone Movielens Report}
\author{Azamat Kurbanaev}
\date{2025-03-01}

@@

\begin{verbatim}
## Loading required package: pacman
\end{verbatim}

\begin{verbatim}
## Warning: package 'pacman' was built under R version 4.4.2
\end{verbatim}

@@

\includegraphics{capstone-movielens-report.draft3_files/figure-latex/unnamed-chunk-10-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft3_files/figure-latex/unnamed-chunk-11-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft3_files/figure-latex/unnamed-chunk-14-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft3_files/figure-latex/unnamed-chunk-15-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft3_files/figure-latex/unnamed-chunk-18-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft3_files/figure-latex/unnamed-chunk-19-1.pdf}

@@

\includegraphics[width=0.4\linewidth]{capstone-movielens-report.draft3_files/figure-latex/sparsity-of-movie-recs-1}

@@

\includegraphics{capstone-movielens-report.draft3_files/figure-latex/movie-id-and-user-hists-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft3_files/figure-latex/unnamed-chunk-32-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft3_files/figure-latex/unnamed-chunk-33-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft3_files/figure-latex/unnamed-chunk-39-1.pdf}
Now we can determine the minimal \emph{RMSE}:

@@

\includegraphics{capstone-movielens-report.draft3_files/figure-latex/unnamed-chunk-45-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft3_files/figure-latex/unnamed-chunk-46-1.pdf}
Below are worst and best ratings categories:

@@

\includegraphics{capstone-movielens-report.draft3_files/figure-latex/unnamed-chunk-48-1.pdf}

@@

To account for \emph{genre effects} we will use the model suggested in
the
\href{https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html\#exercises}{Section
23.7: Exercises} of the \emph{Chapter ``23 Regularization'' of the
Course Textbook}\autocite{IDS2_23-7}:

\[
Y_{i,j} = \mu + \alpha_i + \beta_j + g_{i,j} + \varepsilon_{i,j}
\]

where \(g_{i,j}\) is an \emph{aggregation function} which is explained
in detail in \emph{Section 22.3: ``Review of Aggregation Functions'' of
``Recommender Systems Handbook''} (\emph{Chapter 22: ``Aggregation of
Preferences in Recommender Systems''}, p.~712)
book\autocite{RRSK_RS_HB}.

In the formula above \(g_{i,j}\) denotes a \emph{genre effect} for
user's \(i\) rating of movie \(j\), so that:

\[
g_{i,j} = \sum_{k=1}^K x_{i,j}^k \gamma_k
\]

with \(x^k_{i,j} = 1\) if \(g_{i,j}\) includes genre \(k\), and
\(x^k_{i,j} = 0\) otherwise.

\[
Y_{i,j} = \mu + \alpha_i + \beta_j + g_{i,j} + f(d_{i,j})
\]

\[
 \sum_{i=1}^{n_i} \left(Y_{i,j} - \mu - \alpha_i\right)
\]

### 2025-05-02 reports/capstone-movielens-report.draft4.tex
Draft version 4 of the RMD report compiled.

@@

\title{Capstone Movielens Report}
\author{Azamat Kurbanaev}
\date{2025-03-01}

@@

\begin{verbatim}
## 
## Attaching package: 'data.table'
## 
## The following objects are masked from 'package:lubridate':
## 
##     hour, isoweek, mday, minute, month, quarter, second, wday, week,
##     yday, year
## 
## The following objects are masked from 'package:dplyr':
## 
##     between, first, last
## 
## The following object is masked from 'package:purrr':
## 
##     transpose
## 
## Loading required package: gtools
\end{verbatim}

@@

@@

\begin{verbatim}
## 
## Attaching package: 'scales'
## 
## The following object is masked from 'package:purrr':
## 
##     discard
## 
## The following object is masked from 'package:readr':
## 
##     col_factor
\end{verbatim}

@@

\includegraphics{capstone-movielens-report.draft3_files/figure-latex/unnamed-chunk-10-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft3_files/figure-latex/unnamed-chunk-11-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft3_files/figure-latex/unnamed-chunk-14-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft3_files/figure-latex/unnamed-chunk-15-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft3_files/figure-latex/unnamed-chunk-18-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft3_files/figure-latex/unnamed-chunk-19-1.pdf}

@@

\includegraphics[width=0.4\linewidth]{capstone-movielens-report.draft3_files/figure-latex/sparsity-of-movie-recs-1}

@@

\includegraphics{capstone-movielens-report.draft3_files/figure-latex/movie-id-and-user-hists-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft3_files/figure-latex/unnamed-chunk-32-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft3_files/figure-latex/unnamed-chunk-33-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft3_files/figure-latex/unnamed-chunk-39-1.pdf}
Now we can determine the minimal \emph{RMSE}:

@@

\begin{verbatim}
## # A tibble: 15 x 2
##     year   count
##    <int>   <int>
##  1  1995       2
##  2  1996  942772
##  3  1997  414101
##  4  1998  181634
##  5  1999  709893
##  6  2000 1144349
##  7  2001  683355
##  8  2002  524959
##  9  2003  619938
## 10  2004  691429
## 11  2005 1059277
## 12  2006  689315
## 13  2007  629168
## 14  2008  696740
## 15  2009   13123
\end{verbatim}

@@

\includegraphics{capstone-movielens-report.draft3_files/figure-latex/unnamed-chunk-45-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft3_files/figure-latex/unnamed-chunk-46-1.pdf}
Below are worst and best ratings categories:

@@

\includegraphics{capstone-movielens-report.draft3_files/figure-latex/unnamed-chunk-48-1.pdf}

>>>>

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}

@@

\title{Capstone Movielens Report}
\author{Azamat Kurbanaev}
\date{2025-05-03}

@@

\begin{verbatim}
## 
## Attaching package: 'data.table'
## 
## The following objects are masked from 'package:lubridate':
## 
##     hour, isoweek, mday, minute, month, quarter, second, wday, week,
##     yday, year
## 
## The following objects are masked from 'package:dplyr':
## 
##     between, first, last
## 
## The following object is masked from 'package:purrr':
## 
##     transpose
## 
## Loading required package: gridExtra
## 
## Attaching package: 'gridExtra'
## 
## The following object is masked from 'package:dplyr':
## 
##     combine
## 
## Loading required package: logr
\end{verbatim}

\begin{verbatim}
## Warning: package 'logr' was built under R version 4.4.3
\end{verbatim}

\begin{verbatim}
## Loading required package: common
\end{verbatim}

\begin{verbatim}
## Warning: package 'common' was built under R version 4.4.3
\end{verbatim}

\begin{verbatim}
## Loading required package: gtools
\end{verbatim}

@@

\begin{verbatim}
## Loading required package: recommenderlab
\end{verbatim}

\begin{verbatim}
## Warning: package 'recommenderlab' was built under R version 4.4.3
\end{verbatim}

\begin{verbatim}
## Loading required package: Matrix
\end{verbatim}

\begin{verbatim}
## Warning: package 'Matrix' was built under R version 4.4.2
\end{verbatim}

\begin{verbatim}
## 
## Attaching package: 'Matrix'
## 
## The following objects are masked from 'package:tidyr':
## 
##     expand, pack, unpack
## 
## Loading required package: arules
\end{verbatim}

\begin{verbatim}
## Warning: package 'arules' was built under R version 4.4.3
\end{verbatim}

\begin{verbatim}
## 
## Attaching package: 'arules'
## 
## The following object is masked from 'package:dplyr':
## 
##     recode
## 
## The following objects are masked from 'package:base':
## 
##     abbreviate, write
## 
## Loading required package: proxy
## 
## Attaching package: 'proxy'
## 
## The following object is masked from 'package:Matrix':
## 
##     as.matrix
## 
## The following objects are masked from 'package:stats':
## 
##     as.dist, dist
## 
## The following object is masked from 'package:base':
## 
##     as.matrix
## 
## Registered S3 methods overwritten by 'registry':
##   method               from 
##   print.registry_field proxy
##   print.registry_entry proxy
## 
## Attaching package: 'recommenderlab'
## 
## The following objects are masked from 'package:caret':
## 
##     MAE, RMSE
\end{verbatim}

@@

\begin{verbatim}
## 
## Attaching package: 'scales'
## 
## The following object is masked from 'package:purrr':
## 
##     discard
## 
## The following object is masked from 'package:readr':
## 
##     col_factor
## 
## [conflicted] Will prefer base::as.matrix over any other package.
\end{verbatim}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-10-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-11-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-14-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-15-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-18-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-19-1.pdf}

@@

\includegraphics[width=0.4\linewidth]{capstone-movielens-report.draft4_files/figure-latex/sparsity-of-movie-recs-1}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/movie-id-and-user-hists-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-32-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-33-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-39-1.pdf}
Now we can determine the minimal \emph{RMSE}:

@@

\begin{verbatim}
## # A tibble: 15 x 2
##     year   count
##    <dbl>   <int>
##  1  1995       2
##  2  1996  942772
##  3  1997  414101
##  4  1998  181634
##  5  1999  709893
##  6  2000 1144349
##  7  2001  683355
##  8  2002  524959
##  9  2003  619938
## 10  2004  691429
## 11  2005 1059277
## 12  2006  689315
## 13  2007  629168
## 14  2008  696740
## 15  2009   13123
\end{verbatim}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-45-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-46-1.pdf}
Below are worst and best ratings categories:

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-48-1.pdf}
